---
title: "P4 with BMD raw var"
subtitle: "EPI DATA 536"
editor: visual
date-modified: "today"
format: 
  html: 
    link-external-newwindow: true
    toc: true
    embed-resources: true
    self-contained: true
    page-layout: full
  pdf: default 
editor_options: 
  chunk_output_type: console
---

```{r}
# Libraries
library(tidyverse)
library(survey)
library(gtsummary)
library(tableone)
library(here)
library(broom)

#dataset from P3_PartC
nhanes_newvars<- readRDS(here("data", "processed", "nhanes_bmd_pediatric_newvars.rds"))

```

## Examination of Outcome from Part C + Graphical Visualization

```{r}
# setting male analytic sample
subpop_male <- subset(nhanes_newvars, 
                 subpop1 ==1 & RIAGENDR =="Male")

# setting female analytic sample
subpop_female <- subset(nhanes_newvars, subpop1 ==1 & RIAGENDR =="Female")

attach(nhanes_newvars) 

## Primary survey design 
nhanes_design_full <- svydesign(id = ~SDMVPSU, 
                                strata = ~SDMVSTRA, 
                                weights= ~WTMEC6YR, 
                                nest = TRUE, 
                                data = nhanes_newvars)

# Isolating analytic subpop

subpop_design <- subset(nhanes_design_full, subpop1 == 1)

## Mean & Central Tendency

#Estimate the mean and central tendency of BMD 
## correcting for stratified sampling and survey weights 

svymean(~DXDSTBMD, design = subpop_design)

#Estimating the mean and central tendency of BMD by screen time quartiles 
svyby(formula = ~DXDSTBMD, 
      by = ~factor(screen_quart), 
      design = subpop_design, 
      FUN = svymean, 
      keep.var = TRUE)

#Estimating the mean and central tendency of BMD by screen time quartiles & SEX 
svyby(formula = ~DXDSTBMD, 
      by = ~factor(screen_quart), 
      design = subset(subpop_design, RIAGENDR =="Male"), # by sex male 
      FUN = svymean, 
      keep.var = TRUE)

svyby(formula = ~DXDSTBMD, 
      by = ~factor(screen_quart), 
      design = subset(subpop_design, RIAGENDR =="Female"),# by sex female 
      FUN = svymean, 
      keep.var = TRUE)

# weighted histogram of BMD adjusted z-score
## weighted histogram of analytic sample

svyhist(~DXDSTBMD, design = subpop_design, breaks = 30)

# weighted histogram by male sex
svyhist(~DXDSTBMD, design = subset(subpop_design, RIAGENDR =="Male"), breaks = 30)

# weighted histogram by female sex
svyhist(~DXDSTBMD, design = subset(subpop_design, RIAGENDR =="Female"), breaks = 30)

# weighted boxplot of BMD adjusted z-score
## weighted boxplot
svyboxplot(DXDSTBMD~1, design = subpop_design)

# weighted boxplot by male sex
svyboxplot(DXDSTBMD~1, design = subset(subpop_design, RIAGENDR =="Male"))

# weighted boxplot by female sex
svyboxplot(DXDSTBMD~1, design = subset(subpop_design, RIAGENDR =="Female"))

```

## Visualization of Exposure and Outcome with Linearity & Normality Evaluation
```{r}
## Visualization of Linearity in Exposure and Outcome ##
# Among male subpop
ggplot(subpop_male, 
       aes(x = as.numeric(screen_quart), y = DXDSTBMD)) +
  geom_jitter(size = 1, alpha = .6, width = 0.2) +
  stat_summary(fun = mean, geom = "point", size = 8, shape = 18) +
  geom_smooth(method = lm, se = FALSE) +
  labs(x = "Screen time Quartiles", 
       y = "BMD Raw Score",
       title = "BMD Raw Score vs. screen time quartiles among 8-17 y.o males",
       caption = "Diamonds = screen time level averages") +
  theme(axis.title = element_text(size = 10), 
        axis.text = element_text(size = 10), 
        title = element_text(size = 10), 
        axis.text.x=element_text(angle = 10, vjust = 1, hjust=1))

# Among female subpop
ggplot(subpop_female, 
       aes(x = as.numeric(screen_quart), y = DXDSTBMD)) +
  geom_jitter(size = 1, alpha = .6, width = 0.2) +
  stat_summary(fun = mean, geom = "point", size = 8, shape = 18) +
  geom_smooth(method = lm, se = FALSE) +
  labs(x = "Screen time Quartiles", 
       y = "BMD Raw Score",
       title = "BMD Raw Score vs. screen time quartiles among 8-17 y.o females",
       caption = "Diamonds = screen time level averages") +
  theme(axis.title = element_text(size = 10), 
        axis.text = element_text(size = 10), 
        title = element_text(size = 10), 
        axis.text.x=element_text(angle = 10, vjust = 1, hjust=1))

## Normality##
##Evaluation of distribution of outcome with QQ plot among male subpop
# linear regression
model_male <- svyglm(DXDSTBMD~screen_quart, 
                     design = subset(subpop_design, RIAGENDR =="Male"))

#augment to extract residuals (.resid)
aug_male <- augment(model_male)
glimpse(aug_male)

#creating QQ plot
ggplot(aug_male, 
       aes(sample = .resid)) + 
  stat_qq() + 
  stat_qq_line()

##Evaluation of distribution of outcome with QQ plot among female subpop
# linear regression
model_female <- svyglm(DXDSTBMD~screen_quart, 
                     design = subset(subpop_design, RIAGENDR =="Female"))

#augment to extract residuals (.resid)
aug_female <- augment(model_female)
glimpse(aug_female)

#creating QQ plot
ggplot(aug_female, 
       aes(sample = .resid)) + 
  stat_qq() + 
  stat_qq_line()

```
``

## Visualization of Linearity of other Continuous Independent Variables
### Age as a Continuous Variable
```{r}
## Age as a Continuous Variable ##
# creating a scatterplot in male subpop
nhanes_newvars%>% 
  filter(subpop1==1 & RIAGENDR =="Male")  %>% 
  ggplot(aes(x = RIDAGEYR, 
             y = DXDSTBMD)) + 
  geom_point(position = position_jitter(width = 0.3, height = 0.1)) + 
  labs(title = "Scatterplot of Age and BMD Raw Score", 
      y = "BMD Raw Score")

# overlaying a LOESS (locally weighted regression) plot
nhanes_newvars%>% 
  filter(subpop1==1 & RIAGENDR =="Male")  %>% 
  ggplot(aes(x = RIDAGEYR, 
             y = DXDSTBMD)) +
  geom_point() + 
  geom_smooth(method = "loess", 
              span = 0.75, 
              se = F, 
              color = "green") +
  geom_smooth(method = "loess", 
              span = 0.25, 
              se = F, 
              color = "red", 
              linetype = 2) + 
  theme_classic() 

# Fit the crude model with the linear age term and output the linear predictor
## Using survey specifications coded above with same analytic sample
## run the model among males
crude_age_model <- svyglm(DXDSTBMD~RIDAGEYR, 
                     design = subset(subpop_design, RIAGENDR =="Male"))

summary(crude_age_model)

# Get fitted values
aug_age <- augment(crude_age_model)

# plot the model-predicted linear outcome versus age
ggplot(data = aug_age, 
       aes(x = RIDAGEYR, 
           y = .fitted)) + 
  geom_smooth() + 
  labs(x = "Age at Screening", 
       y = "Value of the Linear Predictor", 
       title = "Linear Model") + 
  theme_classic()

## Higher order predictors
# find mean in subpop
svymean(~RIDAGEYR, design = subpop_design) # weighted mean = 12.482

# create centered variables
nhanes_newvars <- nhanes_newvars %>% 
  mutate(agec = RIDAGEYR - 12.482, 
         agec_sq = agec^2, 
         agec_cu = agec^3)

## Quadratic model
## re-run survey specifications
nhanes_design_quad <- svydesign(
  id = ~SDMVPSU,
  strata = ~SDMVSTRA,
  weights = ~WTMEC6YR,
  nest = TRUE,
  data = nhanes_newvars)

# setting analytic sample
subpop_quad <- subset(nhanes_design_quad, 
                 subpop1 ==1)

## run the model
crude_model_quad <- svyglm(DXDSTBMD~agec+agec_sq, 
                      design = subset(subpop_quad, RIAGENDR =="Male"))

summary(crude_model_quad)

# Get fitted values
model_aug_quad <- augment(crude_model_quad)

ggplot(data = model_aug_quad, 
       aes(x = agec + 12.482, 
           y = .fitted)) + 
  geom_smooth() + 
  labs(x = "Age at Screening", 
       y = "Value of the Linear Predictor", 
       title = "Quadratic Model") + 
  theme_classic()

## Cubic Model
## run the model
crude_model_cube <- svyglm(DXDSTBMD~agec+agec_sq+agec_cu, 
                       design = subset(subpop_quad, RIAGENDR =="Male"))

summary(crude_model_cube)

# Get fitted values
model_aug_cube <- augment(crude_model_cube)

ggplot(data = model_aug_cube, 
       aes(x = agec + 12.482, 
           y = .fitted)) + 
  geom_smooth() + 
  labs(x = "Age at Screening", 
       y = "Value of the Linear Predictor", 
       title = "Cubic Model") + 
  theme_classic()

library(describedata)
gladder(subpop_male$RIDAGEYR) # looking to see if age should be transformed 
# to another power not evaluated. Hard to tell 
```

### Poverty Income Ratio as a Continuous Variable
```{r}
###### Poverty Income Ratio (PIR) as a Continuous Variable #####
# creating a scatterplot in male subpop
nhanes_newvars%>% 
  filter(subpop1==1 & RIAGENDR =="Male")  %>% 
  ggplot(aes(x = INDFMPIR, 
             y = DXDSTBMD)) + 
  geom_point(position = position_jitter(width = 0.3, height = 0.1)) + 
  labs(title = "Scatterplot of PIR and BMD Raw Score", 
      y = "BMD Raw Score", 
      x  = "Poverty Income Ratio (continuous)")

# overlaying a LOESS (locally weighted regression) plot
nhanes_newvars%>% 
  filter(subpop1==1 & RIAGENDR =="Male")  %>% 
  ggplot(aes(x = DXDSTBMD, 
             y = INDFMPIR)) +
  geom_point() + 
  geom_smooth(method = "loess", 
              span = 0.75, 
              se = F, 
              color = "green") +
  geom_smooth(method = "loess", 
              span = 0.25, 
              se = F, 
              color = "red", 
              linetype = 2) + 
  theme_classic() 

# Fit the crude model with the linear age term and output the linear predictor
## Using survey specifications coded above with same analytic sample
## run the model among males
crude_pir_model <- svyglm(DXDSTBMD~INDFMPIR, 
                     design = subset(subpop_design, RIAGENDR =="Male"))

summary(crude_pir_model)

# Get fitted values
aug_pir <- augment(crude_pir_model)

# plot the model-predicted linear outcome versus age
ggplot(data = aug_pir, 
       aes(x = INDFMPIR, 
           y = .fitted)) + 
  geom_smooth() + 
  labs(x = "Poverty Income Ratio at Screening", 
       y = "Value of the Linear Predictor", 
       title = "Linear Model") + 
  theme_classic()

## Create higher order predictors
nhanes_newvars <- nhanes_newvars %>% 
  mutate(pir = INDFMPIR, 
         pir_sq = pir^2, 
         pir_cu = pir^3)

## Quadratic model
## re-run survey specifications
nhanes_design_quad_pir <- svydesign(
  id = ~SDMVPSU,
  strata = ~SDMVSTRA,
  weights = ~WTMEC6YR,
  nest = TRUE,
  data = nhanes_newvars)

# setting analytic sample
subpop_quad_pir <- subset(nhanes_design_quad_pir, 
                 subpop1 ==1)

## run the model
crude_model_quad_pir <- svyglm(DXDSTBMD~pir+pir_sq, 
                      design = subset(subpop_quad_pir, RIAGENDR =="Male"))

summary(crude_model_quad_pir)

# Get fitted values
model_aug_quad_pir <- augment(crude_model_quad_pir)

ggplot(data = model_aug_quad_pir, 
       aes(x = pir, 
           y = .fitted)) + 
  geom_smooth() + 
  labs(x = "Poverty Income ratio at Screening", 
       y = "Value of the Linear Predictor", 
       title = "Quadratic Model") + 
  theme_classic()

## Cubic Model
## run the model
crude_model_cube_pir <- svyglm(DXDSTBMD~pir+pir_sq+pir_cu, 
                       design = subset(subpop_quad_pir, RIAGENDR =="Male"))

summary(crude_model_cube_pir)

# Get fitted values
model_aug_cube_pir <- augment(crude_model_cube_pir)

ggplot(data = model_aug_cube_pir, 
       aes(x = pir, 
           y = .fitted)) + 
  geom_smooth() + 
  labs(x = "Poverty Income Ratio at Screening", 
       y = "Value of the Linear Predictor", 
       title = "Cubic Model") + 
  theme_classic()

pir_clean <- subpop_male %>% 
  filter(INDFMPIR > 0 & !is.na(INDFMPIR)) %>% 
  pull(INDFMPIR)

gladder(pir_clean) # looking to see if PIR should be transformed 
# to another power not evaluated. The best one appears to be "identity". 

``` 

## Estimate of the Crude Measure of Association

```{r}
## Output of simple linear regression model with survey weights for categorical predictor
# among males
summary(model_male)

# among females
summary(model_female)

```

## Estimate the Fully Adjusted Measure of Association
```{r}
# need to relevel the race/ethnicity variable
levels(nhanes_newvars$RIDRETH3)
class(nhanes_newvars$RIDRETH3)

nhanes_newvars <- nhanes_newvars |> 
  mutate(
    race_eth = fct_relevel(RIDRETH3, "Non-Hispanic White")
  )
  
# checking
levels(nhanes_newvars$race_eth)

## resetting the primary survey design 
nhanes_design_full <- svydesign(id = ~SDMVPSU, 
                                strata = ~SDMVSTRA, 
                                weights= ~WTMEC6YR, 
                                nest = TRUE, 
                                data = nhanes_newvars)

# Isolating analytic subpop

subpop_design <- subset(nhanes_design_full, subpop1 == 1)

## Fully adjusted model
# creating the full model 
full_adjust_model <- svyglm(DXDSTBMD~factor(screen_quart) + RIDAGEYR + factor(parent_edu) + factor(race_eth) + INDFMPIR + BMXHT, 
                            design = subpop_design)


summary(full_adjust_model)

# creating the full model among males
full_male_model <- svyglm(DXDSTBMD~factor(screen_quart) + RIDAGEYR + factor(parent_edu) + factor(race_eth) + INDFMPIR + BMXHT, 
                            design = subset(subpop_design, RIAGENDR =="Male"))

summary(full_male_model)

# creating the full model among females
full_female_model <- svyglm(DXDSTBMD~factor(screen_quart) + RIDAGEYR + factor(parent_edu) + factor(race_eth) + INDFMPIR + BMXHT, 
                            design = subset(subpop_design, RIAGENDR =="Female"))

summary(full_female_model)


```

## Tables of Models
```{r}
library(gt)
## Among males
tbl_regression(model_male, label = list(screen_quart = "Screen Quartile"), exponentiate = FALSE, 
               conf.level = 0.95, conf.int = TRUE) |> as_gt() |> 
  tab_header(title = "Crude Model- Males")

tbl_regression(full_male_model, 
               exponentiate = FALSE, 
               include = "factor(screen_quart)", 
               label = list("factor(screen_quart)" = "Screen Quartile"), 
               conf.level = 0.95, 
               conf.int = TRUE) |> as_gt() |> 
  tab_header(title = "Fully Adjusted Model- Males") |> tab_footnote(
    footnote = "Adjusted for child age, parent education, child race/ethnicity, family poverty income ratio, and child height"
  )
               
## Among females
tbl_regression(model_female, label = list(screen_quart = "Screen Quartile"), 
               exponentiate = FALSE, 
               conf.level = 0.95, 
               conf.int = TRUE) |> 
  as_gt() |> 
  tab_header(title = "Crude Model- Females")

tbl_regression(full_female_model, 
               exponentiate = FALSE, 
               include = "factor(screen_quart)", 
               label = list("factor(screen_quart)" = "Screen Quartile"),  
               conf.level = 0.95, 
               conf.int = TRUE) |> 
  as_gt() |> 
  tab_header(title = "Fully Adjusted Model- Females") |> tab_footnote(
    footnote = "Adjusted for child age, parent education, child race/ethnicity, family poverty income ratio, and child height"
  )

```



## Evaluate Effect Measure Modification