---
title: "Part C"
format: html
editor: visual
---

```{r}
# Libraries
library(tidyverse)
library(survey)
library(gtsummary)
library(tableone)
library(here)

# Dataset 
nhanes_clean <- readRDS(here("data", "processed", "nhanes_bmd_pediatric_master.rds"))

# Verification
cat("--- DATASET VERIFICATION ---\n")
cat("Total Raw Sample Loaded (N): ", nrow(nhanes_clean), "\n")
cat("Analytic Subpopulation (subpop1 == 1): ", sum(nhanes_clean$subpop1 == 1, na.rm = TRUE), "\n")
```

```{r}
## Calculated weighted quantiles for exposure
# Temporary survey design to extract weighted quantiles within subpop
temp_design <- svydesign(
  id = ~SDMVPSU,
  strata = ~SDMVSTRA,
  weights = ~WTMEC6YR,
  nest = TRUE,
  data = nhanes_clean) 

#set subpop equal to our analytic sample
subpop_temp <- subset(temp_design, subpop1 ==1)

# Extract weighted 25th, 50th, and 75th percentiles of total screen time
screen_quantiles <- svyquantile(
  ~total_screen_time,
  design = subpop_temp,
  quantiles = c(0.25, 0.50, 0.75),
  na.rm = TRUE
)


# Cutpoints
q25 <- as.numeric(screen_quantiles$total_screen_time[1, "quantile"])
q50 <- as.numeric(screen_quantiles$total_screen_time[2, "quantile"])
q75 <- as.numeric(screen_quantiles$total_screen_time[3, "quantile"])

print(screen_quantiles)
```

```{r}
#create new var for screen time quartiles and set as factor
nhanes_newvars <- nhanes_clean %>%
  mutate(
    screen_quart = case_when(
      total_screen_time <= q25 ~ "Q1",
      total_screen_time >q25 & total_screen_time <= q50 ~ "Q2",
      total_screen_time >q50 & total_screen_time <= q75 ~ "Q3",
      total_screen_time >q75 ~ "Q4",
    .default = NA_character_) %>%
      factor(levels = c("Q1", "Q2", "Q3", "Q4")))

#unweighted sample size per quartile, restricted to subpop1: >4,600 in each
table(nhanes_newvars$screen_quart[nhanes_newvars$subpop1 == 1], useNA = "always")

#create new var for FPL categories
nhanes_newvars <- nhanes_newvars %>%
  mutate(
    fpl_cat = case_when(
      INDFMPIR < 1.30 ~ "Low Income (< 130% PIR)",
      INDFMPIR >= 1.30 & INDFMPIR <= 3.50 ~ "Middle Income (130-350% PIR)",
      INDFMPIR > 3.50 ~ "High Income (> 350% PIR)",
      .default = NA_character_
    ) %>% factor(levels = c("Low Income (< 130% PIR)", "Middle Income (130-350% PIR)", "High Income (> 350% PIR)")))

# check to make sure no missing values for z scores 
sum(nhanes_clean$subpop1 == 1 & is.na(nhanes_clean$bmd_zscore_adj))
#Create categorical version of TBLH BMD z-scores for prevalence estimates
nhanes_newvars <- nhanes_newvars %>%
  mutate(
    bmd_z_cat = case_when(
      bmd_zscore_adj <= -2.0 ~ "Low TBLH BMD",
      bmd_zscore_adj > -2.0 ~ "Normal TBLH BMD",
      .default = NA_character_
    ) %>% factor(levels = c("Normal TBLH BMD", "Low TBLH BMD")))

#confirm factors in nhanes_newvars
nhanes_newvars %>% 
  select(RIAGENDR, RIDRETH3, fpl_cat, screen_quart, bmd_z_cat, FSDCH) %>% 
  glimpse()
```

```{r}
## Primary survey design
nhanes_design_full <- svydesign(
  id = ~SDMVPSU,
  strata = ~SDMVSTRA,
  weights = ~WTMEC6YR,
  nest = TRUE,
  data = nhanes_newvars)

# Isolating analytic subpop
subpop_design <- subset(nhanes_design_full, subpop1 == 1)
```

```{r}
## Prevalence Estiimate ##
#Estimate prevalence of low BMD (defined as z score less than -2.0) using weighted estimates
svyciprop(~I(bmd_z_cat == "Low TBLH BMD"),
          design=subpop_design,
          method = "logit",
          level = 0.95)
```

```{r}
## Table 1 based on course suggestion ##
#load libraries
library(gtsummary)
library(rstatix)
library(jtools)
library(tidyverse)
library(dplyr)
library(here)
library(haven)
library(psych)
library(janitor)
library(survey)
library(Hmisc)

#recode names of values for FSDCH variable
table(nhanes_newvars$FSDCH, useNA = "ifany")
nhanes_newvars <- nhanes_newvars %>%
  mutate(
    FSDCH = fct_recode(
      FSDCH,
    "High/Marginal Food Security"  = "CH full or marginal food security: 0",
    "Marginal Food Security"      = "CH marginal food security: 1",
    "Low Food Security"           = "CH low food security: 2-4",
    "Very Low Food Security"      = "CH very low food security: 5-8"))

#set levels for FSDCH - high/marginal as ref
nhanes_newvars <- nhanes_newvars %>%
  mutate(
    FSDCH = factor(
      FSDCH,
      levels = c(
        "High/Marginal Food Security",
        "Marginal Food Security",
        "Low Food Security",
        "Very Low Food Security")))

# creating labels for variable names to display on Table 1
# cannot label the stratification lables or else the code won't work for the Table
label(nhanes_newvars$RIDRETH3) <- "Race/Ethnicity"
label(nhanes_newvars$fpl_cat) <- "Household Poverty Income Ratio (PIR)"
label(nhanes_newvars$parent_edu) <- "Parent Education"
label(nhanes_newvars$RIDAGEYR) <- "Age"
label(nhanes_newvars$FSDCH) <- "Food Security Status"


#check results
tabyl(nhanes_newvars, FSDCH)

#establish survey design
nhanes_design_full <- svydesign(
  id = ~SDMVPSU,
  strata = ~SDMVSTRA,
  weights = ~WTMEC6YR,
  nest = TRUE,
  data = nhanes_newvars)

# Isolating analytic subpop
subpop_design <- subset(nhanes_design_full, subpop1 == 1)

# Subset designs by sex
design_males   <- subset(subpop_design, RIAGENDR %in% c("Male", "male", 1))
design_females <- subset(subpop_design, RIAGENDR %in% c("Female", "female", 2))

#Build table 1 - males
library(gtsummary)
library(dplyr)

t1_male <- tbl_svysummary(
  data = design_males,
  by = screen_quart,
  include = c(RIDAGEYR, RIDRETH3, fpl_cat, parent_edu, FSDCH),
  missing = "no",
  statistic = list(
    all_continuous() ~ "{mean} ({mean.std.error})", # using Mean and SE for all contin vars
    all_categorical() ~ "{n_unweighted} ({p}%)") # using unweighted counts and wtd % for all cat vars
  ) %>%
  add_n("{N_miss_unweighted} ({p_miss})") %>% # adding a column for Missing unwtd counts and wtdN(these are not variables; they are part of the function) %
  modify_header(n ~ "**Missing n (%)**") %>% #creating the label for the header of Missing
  modify_spanning_header(all_stat_cols() ~ "**TBLH BMD Z-score Quartile**") %>%  #labeling the stratification var
  modify_header(stat_1 ~ "**Q1**<br>N = {n_unweighted}") %>% #labeling strat columns & getting unwted n
  modify_header(stat_2 ~ "**Q2**<br>N = {n_unweighted}") %>%
  modify_header(stat_3 ~ "**Q3**<br>N = {n_unweighted}") %>%
  modify_header(stat_4 ~ "**Q4**<br>N = {n_unweighted}") %>%
  modify_caption(caption = "Table 1. NHANES 2013-2018 distribution of characteristics among U.S male children 8 to 17 years of age (n = 2,044) according to quartiles of average daily screen time ") %>%
  #remove default footnote and superscript
modify_footnote(all_stat_cols() ~ NA) %>%
  as_gt() %>%
  gt::tab_source_note(
    source_note = "Weighted percentages and unweighted sample counts shown. Abbreviations: NHANES, National Health and Nutrition Examination Survey; SE, Standard Error; PIR, Poverty Income Ratio; TBLH, Total Body Less Head; BMD, Bone Mineral Density."
  )

t1_male
```

```{r}
#Build table 1 - females

t1_female <- tbl_svysummary(
  data = design_females,
  by = screen_quart,
  include = c(RIDAGEYR, RIDRETH3, fpl_cat, parent_edu, FSDCH),
  missing = "no",
  statistic = list(
    all_continuous() ~ "{mean} ({mean.std.error})", # using Mean and SE for all contin vars
    all_categorical() ~ "{n_unweighted} ({p}%)") # using unweighted counts and wtd % for all cat vars
  ) %>%
  add_n("{N_miss_unweighted} ({p_miss})") %>% # adding a column for Missing unwtd counts and wtdN(these are not variables; they are part of the function) %
  modify_header(n ~ "**Missing n (%)**") %>% #creating the label for the header of Missing
  modify_spanning_header(all_stat_cols() ~ "**TBLH BMD Z-score Quartile**") %>%  #labeling the stratification var
  modify_header(stat_1 ~ "**Q1**<br>N = {n_unweighted}") %>% #labeling strat columns & getting unwted n
  modify_header(stat_2 ~ "**Q2**<br>N = {n_unweighted}") %>%
  modify_header(stat_3 ~ "**Q3**<br>N = {n_unweighted}") %>%
  modify_header(stat_4 ~ "**Q4**<br>N = {n_unweighted}") %>%
  modify_caption(caption = "Table 1. NHANES 2013-2018 distribution of characteristics among U.S female children 8 to 17 years of age (n = 1,953) according to quartiles of average daily screen time ") %>%
  #remove default footnote and superscript
modify_footnote(all_stat_cols() ~ NA) %>%
  as_gt() %>%
  gt::tab_source_note(
    source_note = "Weighted percentages and unweighted sample counts shown. Abbreviations: NHANES, National Health and Nutrition Examination Survey; SE, Standard Error; PIR, Poverty Income Ratio; TBLH, Total Body Less Head; BMD, Bone Mineral Density."
  )

t1_female
```

```{r}
## Save table 1 ##
gt::gtsave(t1_male, filename = "Table1_Males.docx") #save male as word doc
gt::gtsave(t1_female, filename = "Table1_Females.docx") #save male as word doc

gt::gtsave(t1_male, filename = "Table1_Males.png") #save male as png
gt::gtsave(t1_female, filename = "Table1_Females.png") #save female as png
```

Old code for making table 1. Can prob delete, but saving just in case for now.

library(gtsummary) library(survey)

# Subset designs by sex

design_males \<- subset(subpop_design, RIAGENDR %in% c("Male", "male", 1)) design_females \<- subset(subpop_design, RIAGENDR %in% c("Female", "female", 2))

# Table 1 Variables

tb1_vars \<- c("RIDAGEYR", "RIDRETH3", "parent_edu", "fpl_cat", "FSDCH")

#Table 1 - MALES table1_males \<- design_males %\>% tbl_svysummary( by = screen_quart, include = all_of(tb1_vars), statistic = list( all_continuous() \~ "{mean} ({sd})", all_categorical() \~ "{n_unweighted} ({p}%)" ), digits = all_categorical() \~ c(0, 1) ) %\>% add_overall(col_label = "**Overall** (N = {N_unweighted})") %\>% bold_labels()

table1_males

#Table 1 - FEMALES table1_females \<- design_females %\>% tbl_svysummary( by = screen_quart, include = all_of(tb1_vars), statistic = list( all_continuous() \~ "{mean} ({sd})", all_categorical() \~ "{n_unweighted} ({p}%)" ), digits = all_categorical() \~ c(0, 1) ) %\>% add_overall(col_label = "**Overall** (N = {N_unweighted})") %\>% bold_labels()

table1_females

#create subset for males and females (unweighted) unweighted_males \<- nhanes_newvars %\>% filter(subpop1 == 1 & RIAGENDR == "Male")

unweighted_females \<- nhanes_newvars %\>% filter(subpop1 == 1 & RIAGENDR == "Female")

nrow(unweighted_males) #verify numbers nrow(unweighted_females)

#Unweighted counts unweighted_males %\>% freqtables::freq_table(screen_quart, RIDRETH3) %\>% select( screen_quart = row_cat, race_ethnicity = col_cat, n, percent = percent_row )
