---
title: "P4 with BMD raw var"
subtitle: "EPI DATA 536"
editor: visual
date-modified: "today"
format: 
  html: 
    link-external-newwindow: true
    toc: true
    embed-resources: true
    self-contained: true
    page-layout: full
  pdf: default 
editor_options: 
  chunk_output_type: console
---

```{r}
# Libraries
library(tidyverse)
library(survey)
library(gtsummary)
library(tableone)
library(here)
library(broom)

#dataset from P3_PartC
nhanes_screen <- readRDS("C:/Users/anabe/OneDrive/Desktop/New_BMD-Project/data/processed/nhanes_bmd_pediatric_newvars_rawoutcome1.rds")
```

## Examination of Outcome from Part C + Graphical Visualization

```{r}
# setting male analytic sample
subpop_male <- subset(nhanes_screen, 
                 subpop1 ==1 & RIAGENDR =="Male")

# setting female analytic sample
subpop_female <- subset(nhanes_screen, subpop1 ==1 & RIAGENDR =="Female")

attach(nhanes_screen) 

## Primary survey design 
nhanes_design_full <- svydesign(id = ~SDMVPSU, 
                                strata = ~SDMVSTRA, 
                                weights= ~WTMEC6YR, 
                                nest = TRUE, 
                                data = nhanes_screen)

# Isolating general analytic subpop

subpop_design <- subset(nhanes_design_full, subpop1 == 1)

## Mean & Central Tendency

#Estimate the mean and central tendency of BMD 
## correcting for stratified sampling and survey weights 

svymean(~DXDSTBMD, design = subpop_design)

#Estimating the mean and central tendency of BMD by screen time quartiles 
svyby(formula = ~DXDSTBMD, 
      by = ~factor(screen_quart), 
      design = subpop_design, 
      FUN = svymean, 
      keep.var = TRUE)

#Estimating the mean and central tendency of BMD by screen time quartiles & SEX 
svyby(formula = ~DXDSTBMD, 
      by = ~factor(screen_quart), 
      design = subset(subpop_design, RIAGENDR =="Male"), # by sex male 
      FUN = svymean, 
      keep.var = TRUE)

svyby(formula = ~DXDSTBMD, 
      by = ~factor(screen_quart), 
      design = subset(subpop_design, RIAGENDR =="Female"),# by sex female 
      FUN = svymean, 
      keep.var = TRUE)

# weighted histogram of BMD adjusted z-score
## weighted histogram of analytic sample

svyhist(~DXDSTBMD, design = subpop_design, breaks = 30)

# weighted histogram by male sex
svyhist(~DXDSTBMD, design = subset(subpop_design, RIAGENDR =="Male"), breaks = 30)

# weighted histogram by female sex
svyhist(~DXDSTBMD, design = subset(subpop_design, RIAGENDR =="Female"), breaks = 30)

# weighted boxplot of BMD adjusted z-score
## weighted boxplot
svyboxplot(DXDSTBMD~1, design = subpop_design)

# weighted boxplot by male sex
svyboxplot(DXDSTBMD~1, design = subset(subpop_design, RIAGENDR =="Male"))

# weighted boxplot by female sex
svyboxplot(DXDSTBMD~1, design = subset(subpop_design, RIAGENDR =="Female"))

```

## Visualization of Exposure and Outcome with Linearity & Normality Evaluation

```{r}
## Visualization of Linearity in Exposure and Outcome ##
# Among male subpop
ggplot(subpop_male, 
       aes(x = as.numeric(screen_quart), y = DXDSTBMD)) +
  geom_jitter(size = 1, alpha = .6, width = 0.2) +
  stat_summary(fun = mean, geom = "point", size = 8, shape = 18) +
  geom_smooth(method = lm, se = FALSE) +
  labs(x = "Screen time Quartiles", 
       y = "BMD Raw Score",
       title = "BMD Raw Score vs. screen time quartiles among 8-17 y.o males",
       caption = "Diamonds = screen time level averages") +
  theme(axis.title = element_text(size = 10), 
        axis.text = element_text(size = 10), 
        title = element_text(size = 10), 
        axis.text.x=element_text(angle = 10, vjust = 1, hjust=1))

# Among female subpop
ggplot(subpop_female, 
       aes(x = as.numeric(screen_quart), y = DXDSTBMD)) +
  geom_jitter(size = 1, alpha = .6, width = 0.2) +
  stat_summary(fun = mean, geom = "point", size = 8, shape = 18) +
  geom_smooth(method = lm, se = FALSE) +
  labs(x = "Screen time Quartiles", 
       y = "BMD Raw Score",
       title = "BMD Raw Score vs. screen time quartiles among 8-17 y.o females",
       caption = "Diamonds = screen time level averages") +
  theme(axis.title = element_text(size = 10), 
        axis.text = element_text(size = 10), 
        title = element_text(size = 10), 
        axis.text.x=element_text(angle = 10, vjust = 1, hjust=1))

## Normality##
##Evaluation of distribution of outcome with QQ plot among male subpop
# linear regression
model_male <- svyglm(DXDSTBMD~factor(screen_quart), 
                     design = subset(subpop_design, RIAGENDR =="Male"))

#augment to extract residuals (.resid)
aug_male <- augment(model_male)
glimpse(aug_male)

#creating QQ plot
ggplot(aug_male, 
       aes(sample = .resid)) + 
  stat_qq() + 
  stat_qq_line()

##Evaluation of distribution of outcome with QQ plot among female subpop
# linear regression
model_female <- svyglm(DXDSTBMD~factor(screen_quart), 
                     design = subset(subpop_design, RIAGENDR =="Female"))

#augment to extract residuals (.resid)
aug_female <- augment(model_female)
glimpse(aug_female)

#creating QQ plot
ggplot(aug_female, 
       aes(sample = .resid)) + 
  stat_qq() + 
  stat_qq_line()

```

## Visualization of Linearity of other Continuous Independent Variables

### Age as a Continuous Variable

```{r}
## Age as a Continuous Variable ##
# creating a scatterplot in male subpop
nhanes_screen%>% 
  filter(subpop1==1 & RIAGENDR =="Male")  %>% 
  ggplot(aes(x = RIDAGEYR, 
             y = DXDSTBMD)) + 
  geom_point(position = position_jitter(width = 0.3, height = 0.1)) + 
  labs(title = "Scatterplot of Age and BMD Raw Score", 
      y = "BMD Raw Score")

# overlaying a LOESS (locally weighted regression) plot
nhanes_screen%>% 
  filter(subpop1==1 & RIAGENDR =="Male")  %>% 
  ggplot(aes(x = RIDAGEYR, 
             y = DXDSTBMD)) +
  geom_point() + 
  geom_smooth(method = "loess", 
              span = 0.75, 
              se = F, 
              color = "green") +
  geom_smooth(method = "loess", 
              span = 0.25, 
              se = F, 
              color = "red", 
              linetype = 2) + 
  theme_classic() 

# Fit the crude model with the linear age term and output the linear predictor
## Using survey specifications coded above with same analytic sample
## run the model among males
crude_age_model <- svyglm(DXDSTBMD~RIDAGEYR, 
                     design = subset(subpop_design, RIAGENDR =="Male"))

summary(crude_age_model)

# Get fitted values
aug_age <- augment(crude_age_model)

# plot the model-predicted linear outcome versus age
ggplot(data = aug_age, 
       aes(x = RIDAGEYR, 
           y = .fitted)) + 
  geom_smooth() + 
  labs(x = "Age at Screening", 
       y = "Value of the Linear Predictor", 
       title = "Linear Model") + 
  theme_classic()

## Higher order predictors
# find mean in subpop
svymean(~RIDAGEYR, design = subpop_design) # weighted mean = 12.482

# create centered variables
nhanes_screen <- nhanes_screen %>% 
  mutate(agec = RIDAGEYR - 12.482, 
         agec_sq = agec^2, 
         agec_cu = agec^3)

## Quadratic model
## re-run survey specifications
nhanes_design_quad <- svydesign(
  id = ~SDMVPSU,
  strata = ~SDMVSTRA,
  weights = ~WTMEC6YR,
  nest = TRUE,
  data = nhanes_screen)

# setting analytic sample
subpop_quad <- subset(nhanes_design_quad, 
                 subpop1 ==1)

## run the model
crude_model_quad <- svyglm(DXDSTBMD~agec+agec_sq, 
                      design = subset(subpop_quad, RIAGENDR =="Male"))

summary(crude_model_quad)

# Get fitted values
model_aug_quad <- augment(crude_model_quad)

ggplot(data = model_aug_quad, 
       aes(x = agec + 12.482, 
           y = .fitted)) + 
  geom_smooth() + 
  labs(x = "Age at Screening", 
       y = "Value of the Linear Predictor", 
       title = "Quadratic Model") + 
  theme_classic()

## Cubic Model
## run the model
crude_model_cube <- svyglm(DXDSTBMD~agec+agec_sq+agec_cu, 
                       design = subset(subpop_quad, RIAGENDR =="Male"))

summary(crude_model_cube)

# Get fitted values
model_aug_cube <- augment(crude_model_cube)

ggplot(data = model_aug_cube, 
       aes(x = agec + 12.482, 
           y = .fitted)) + 
  geom_smooth() + 
  labs(x = "Age at Screening", 
       y = "Value of the Linear Predictor", 
       title = "Cubic Model") + 
  theme_classic()

library(describedata)
gladder(subpop_male$RIDAGEYR) # looking to see if age should be transformed 
# to another power not evaluated. Hard to tell 
```

### Poverty Income Ratio as a Continuous Variable

```{r}
###### Poverty Income Ratio (PIR) as a Continuous Variable #####
# creating a scatterplot in male subpop
nhanes_screen%>% 
  filter(subpop1==1 & RIAGENDR =="Male")  %>% 
  ggplot(aes(x = INDFMPIR, 
             y = DXDSTBMD)) + 
  geom_point(position = position_jitter(width = 0.3, height = 0.1)) + 
  labs(title = "Scatterplot of PIR and BMD Raw Score", 
      y = "BMD Raw Score", 
      x  = "Poverty Income Ratio (continuous)")

# overlaying a LOESS (locally weighted regression) plot
nhanes_screen%>% 
  filter(subpop1==1 & RIAGENDR =="Male")  %>% 
  ggplot(aes(x = DXDSTBMD, 
             y = INDFMPIR)) +
  geom_point() + 
  geom_smooth(method = "loess", 
              span = 0.75, 
              se = F, 
              color = "green") +
  geom_smooth(method = "loess", 
              span = 0.25, 
              se = F, 
              color = "red", 
              linetype = 2) + 
  theme_classic() 

# Fit the crude model with the linear age term and output the linear predictor
## Using survey specifications coded above with same analytic sample
## run the model among males
crude_pir_model <- svyglm(DXDSTBMD~INDFMPIR, 
                     design = subset(subpop_design, RIAGENDR =="Male"))

summary(crude_pir_model)

# Get fitted values
aug_pir <- augment(crude_pir_model)

# plot the model-predicted linear outcome versus age
ggplot(data = aug_pir, 
       aes(x = INDFMPIR, 
           y = .fitted)) + 
  geom_smooth() + 
  labs(x = "Poverty Income Ratio at Screening", 
       y = "Value of the Linear Predictor", 
       title = "Linear Model") + 
  theme_classic()

## Create higher order predictors
nhanes_screen <- nhanes_screen %>% 
  mutate(pir = INDFMPIR, 
         pir_sq = pir^2, 
         pir_cu = pir^3)

## Quadratic model
## re-run survey specifications
nhanes_design_quad_pir <- svydesign(
  id = ~SDMVPSU,
  strata = ~SDMVSTRA,
  weights = ~WTMEC6YR,
  nest = TRUE,
  data = nhanes_screen)

# setting analytic sample
subpop_quad_pir <- subset(nhanes_design_quad_pir, 
                 subpop1 ==1)

## run the model
crude_model_quad_pir <- svyglm(DXDSTBMD~pir+pir_sq, 
                      design = subset(subpop_quad_pir, RIAGENDR =="Male"))

summary(crude_model_quad_pir)

# Get fitted values
model_aug_quad_pir <- augment(crude_model_quad_pir)

ggplot(data = model_aug_quad_pir, 
       aes(x = pir, 
           y = .fitted)) + 
  geom_smooth() + 
  labs(x = "Poverty Income ratio at Screening", 
       y = "Value of the Linear Predictor", 
       title = "Quadratic Model") + 
  theme_classic()

## Cubic Model
## run the model
crude_model_cube_pir <- svyglm(DXDSTBMD~pir+pir_sq+pir_cu, 
                       design = subset(subpop_quad_pir, RIAGENDR =="Male"))

summary(crude_model_cube_pir)

# Get fitted values
model_aug_cube_pir <- augment(crude_model_cube_pir)

ggplot(data = model_aug_cube_pir, 
       aes(x = pir, 
           y = .fitted)) + 
  geom_smooth() + 
  labs(x = "Poverty Income Ratio at Screening", 
       y = "Value of the Linear Predictor", 
       title = "Cubic Model") + 
  theme_classic()

pir_clean <- subpop_male %>% 
  filter(INDFMPIR > 0 & !is.na(INDFMPIR)) %>% 
  pull(INDFMPIR)

gladder(pir_clean) # looking to see if PIR should be transformed 
# to another power not evaluated. The best one appears to be "identity". 

```

## Estimates of the Crude Measure of Association

```{r}
## Output of simple linear regression model with survey weights for categorical predictor
# among males
summary(model_male)

# among females
summary(model_female)

```

## Estimates of the Fully Adjusted Measure of Association

```{r}
# need to relevel the race/ethnicity variable
levels(nhanes_screen$RIDRETH3)
class(nhanes_screen$RIDRETH3)

nhanes_screen <- nhanes_screen |> 
  mutate(
    race_eth = fct_relevel(RIDRETH3, "Non-Hispanic White")
  )
  
# checking
levels(nhanes_screen$race_eth)

## resetting the primary survey design 
nhanes_design_full <- svydesign(id = ~SDMVPSU, 
                                strata = ~SDMVSTRA, 
                                weights= ~WTMEC6YR, 
                                nest = TRUE, 
                                data = nhanes_screen)

# Isolating analytic subpop

subpop_design <- subset(nhanes_design_full, subpop1 == 1)

## Fully adjusted model
# creating the full model 
full_adjust_model <- svyglm(DXDSTBMD~factor(screen_quart) + RIDAGEYR + factor(parent_edu) + factor(race_eth) + INDFMPIR + BMXHT, 
                            design = subpop_design)


summary(full_adjust_model)

# creating the full model among males
full_male_model <- svyglm(DXDSTBMD~factor(screen_quart) + RIDAGEYR + factor(parent_edu) + factor(race_eth) + INDFMPIR + BMXHT, 
                            design = subset(subpop_design, RIAGENDR =="Male"))

summary(full_male_model)

# creating the full model among females
full_female_model <- svyglm(DXDSTBMD~factor(screen_quart) + RIDAGEYR + factor(parent_edu) + factor(race_eth) + INDFMPIR + BMXHT, 
                            design = subset(subpop_design, RIAGENDR =="Female"))

summary(full_female_model)


```

## Creation of Table 2 and Table 3

### Among Males

```{r}
library(gtsummary)
library(gt)
library(srvyr)
library(dplyr)
library(survey)


# 1. Distribution of Screen Quartile (Unweighted N, Weighted %)
w_pct <- tbl_svysummary(
  data = subset(subpop_design, RIAGENDR == "Male"), 
  include = screen_quart, 
  missing = "no", 
  statistic = list(all_categorical() ~ "{n_unweighted} ({p}%)"), 
  label = list(screen_quart = "Screen Quartile")
)

male_design <- subset(nhanes_design_full, subpop1 == 1 & RIAGENDR == "Male")


# 2. Mean BMD Table (Vertical Layout)
# -------------------------------------------------------------
# Calculate weighted means and SEs by screen_quart
bmd_by_q <- svyby(~DXDSTBMD, ~screen_quart, male_design, svymean, na.rm = TRUE)

# Create a named lookup vector formatted as "Mean (SE)"
bmd_lookup <- setNames(
  sprintf("%.3f (%.3f)", bmd_by_q$DXDSTBMD, bmd_by_q$se),
  as.character(bmd_by_q$screen_quart)
)

# Copy w_pct structure and map the Mean (SE) values onto the quartile levels
m_mean <- w_pct
m_mean$table_body$stat_0 <- ifelse(
  m_mean$table_body$row_type == "level",
  bmd_lookup[as.character(m_mean$table_body$label)],
  NA_character_
)

m_mean
  
# 3. Unadjusted Regression Table
t2 <- tbl_regression(
  model_male, 
  include = dplyr::everything(),
  label = list("factor(screen_quart)" = "Screen Quartile"), 
  exponentiate = FALSE, 
  conf.level = 0.95, 
  conf.int = TRUE
) 

# 4. Adjusted Regression Table
t3 <- tbl_regression(
  full_male_model, 
  exponentiate = FALSE, 
  include = "factor(screen_quart)", 
  label = list("factor(screen_quart)" = "Screen Quartile"), 
  conf.level = 0.95, 
  conf.int = TRUE
)

# Harmonize internal variable names to match w_pct ("screen_quart")
t2$table_body$variable[t2$table_body$variable == "factor(screen_quart)"] <- "screen_quart"
t3$table_body$variable[t3$table_body$variable == "factor(screen_quart)"] <- "screen_quart"

merged_male_table <- tbl_merge(
  tbls = list(w_pct, m_mean, t2, t3),
  tab_spanner = c(
    "**Sample Size**",
    "**Mean BMD (SE)**",
    "**Unadjusted Model**",
    "**Adjusted Model** <sup>2</sup>"
  )
) %>%
  modify_caption(
    caption = "Table 2. Associations beetween daily average screen time and TBLH BMD raw scores in unadjusted and fully adjusted linear regression models among male children aged 8- 17 y.o (n = 2,044)"
  ) %>%
  modify_header(
    all_stat_cols() ~ "**{level}**<br>N = {n_unweighted}"
  ) %>%
  as_gt() %>%
  gt::tab_source_note(
    source_note = "² Fully adjusted models were adjusted for age, race/ethnicity, height (in cm), household reference person's education level, and household income evaluated at the federal poverty level."
  )

merged_male_table


# attempting to save the table 
gt::gtsave(merged_male_table, filename = "Table2_Males.png") #save male as png


```

### Among Females

```{r}
# 1. Distribution of Screen Quartile (Unweighted N, Weighted %)
w_pct <- tbl_svysummary(
  data = subset(subpop_design, RIAGENDR == "Female"), 
  include = screen_quart, 
  missing = "no", 
  statistic = list(all_categorical() ~ "{n_unweighted} ({p}%)"), 
  label = list(screen_quart = "Screen Quartile")
)


female_design <- subset(nhanes_design_full, subpop1 == 1 & RIAGENDR == "Female")

# 2. Mean BMD Table (Vertical Layout)
# -------------------------------------------------------------
# Calculate weighted means and SEs by screen_quart
bmd_by_q <- svyby(~DXDSTBMD, ~screen_quart, female_design, svymean, na.rm = TRUE)

# Create a named lookup vector formatted as "Mean (SE)"
bmd_lookup <- setNames(
  sprintf("%.3f (%.3f)", bmd_by_q$DXDSTBMD, bmd_by_q$se),
  as.character(bmd_by_q$screen_quart)
)

# Copy w_pct structure and map the Mean (SE) values onto the quartile levels
m_mean <- w_pct
m_mean$table_body$stat_0 <- ifelse(
  m_mean$table_body$row_type == "level",
  bmd_lookup[as.character(m_mean$table_body$label)],
  NA_character_
)

m_mean
  
# 3. Unadjusted Regression Table
t2 <- tbl_regression(
  model_female, 
  include = dplyr::everything(),
  label = list("factor(screen_quart)" = "Screen Quartile"), 
  exponentiate = FALSE, 
  conf.level = 0.95, 
  conf.int = TRUE
) 

# 4. Adjusted Regression Table
t3 <- tbl_regression(
  full_female_model, 
  exponentiate = FALSE, 
  include = "factor(screen_quart)", 
  label = list("factor(screen_quart)" = "Screen Quartile"), 
  conf.level = 0.95, 
  conf.int = TRUE
)

# Harmonize internal variable names to match w_pct ("screen_quart")
t2$table_body$variable[t2$table_body$variable == "factor(screen_quart)"] <- "screen_quart"
t3$table_body$variable[t3$table_body$variable == "factor(screen_quart)"] <- "screen_quart"

merged_female_table <- tbl_merge(
  tbls = list(w_pct, m_mean, t2, t3),
  tab_spanner = c(
    "**Sample Size**",
    "**Mean BMD (SE)**",
    "**Unadjusted Model**",
    "**Adjusted Model** <sup>2</sup>"
  )
) %>%
  modify_caption(caption = "Table 3. Associations beetween daily average screen time and TBLH BMD raw scores in unadjusted and fully adjusted linear regression models among female children aged 8- 17 y.o (n = 1,953)") %>%
  modify_header(
    all_stat_cols() ~ "**{level}**<br>N = {n_unweighted}"
  ) |> 
  as_gt() %>%
  gt::tab_source_note(
    source_note = "² Fully adjusted models were adjusted for age, race/ethnicity, height (in cm), household reference person's education level, and household income evaluated at the federal poverty level."
  )

merged_female_table

# attempting to save the table 
gt::gtsave(merged_female_table, filename = "Table2_Females.png") #save male as png


```

## Evaluation of Effect Measure Modification Food Security- Interaction

```{r}
## Evaluating EMM with interaction term in the model
# Among Males
library(jtools)
int_male_model <- svyglm(DXDSTBMD~factor(screen_quart)*factor(fsdch_2cat) + RIDAGEYR + factor(parent_edu) + factor(race_eth) + INDFMPIR + BMXHT, 
                            design = subset(subpop_design, RIAGENDR =="Male"))

# view the model
summ(int_male_model, 
     confint = T, 
     ci.width = 0.95, 
     exp = F)

# Wald Test of Multiplicative Interaction
regTermTest(int_male_model, ~factor(screen_quart):factor(fsdch_2cat), 
            method = "Wald")

# Among Females
int_female_model <- svyglm(DXDSTBMD~factor(screen_quart)*factor(fsdch_2cat) + RIDAGEYR + factor(parent_edu) + factor(race_eth) + INDFMPIR + BMXHT, 
                            design = subset(subpop_design, RIAGENDR =="Female"))

# view the model
summ(int_female_model, 
     confint = T, 
     ci.width = 0.95, 
     exp = F)

# Wald Test of Multiplicative Interaction
regTermTest(int_female_model, ~factor(screen_quart):factor(fsdch_2cat), 
            method = "Wald")

```

## Evaluation of Effect Measure Modification Food Security- Stratification

```{r}
# looking at levels of food security
levels(nhanes_screen$fsdch_2cat)


### using the same survey specifications and analytic sample as in chunk 2
## Among Males with Food Security
reg_male_fs <- svyglm(DXDSTBMD~factor(screen_quart) + RIDAGEYR + factor(parent_edu) + factor(race_eth) + INDFMPIR + BMXHT, 
                          design = subset(subpop_design, RIAGENDR =="Male"
                                          & fsdch_2cat == "High/Marginal Food Security"))

summary(reg_male_fs) # crude values

#confidence interval, counts, p-val, etc 
summ(reg_male_fs, confint = T) 


## Among Males with Food Insecurity
reg_male_fins <- svyglm(DXDSTBMD~factor(screen_quart) + RIDAGEYR + factor(parent_edu) + factor(race_eth) + INDFMPIR + BMXHT, 
                        design = subset(subpop_design, RIAGENDR =="Male"
                                        & fsdch_2cat == "Low/Very Low Food Security"))

summary(reg_male_fins) # crude values

#confidence interval, counts, p-val, etc 
summ(reg_male_fins, confint = T) 

## Among Females with Food Security
reg_female_fs <- svyglm(DXDSTBMD~factor(screen_quart) + RIDAGEYR + factor(parent_edu) + factor(race_eth) + INDFMPIR + BMXHT, 
                        design = subset(subpop_design, RIAGENDR =="Female"
                                        & fsdch_2cat == "High/Marginal Food Security"))

summary(reg_female_fs) # crude values

#confidence interval, counts, p-val, etc 
summ(reg_female_fs, confint = T) 


## Among Females with Food Insecurity
reg_female_fins <- svyglm(DXDSTBMD~factor(screen_quart) + RIDAGEYR + factor(parent_edu) + factor(race_eth) + INDFMPIR + BMXHT, 
                          design = subset(subpop_design, RIAGENDR =="Female"
                                          & fsdch_2cat == "Low/Very Low Food Security"))

summary(reg_female_fins) # crude values

#confidence interval, counts, p-val, etc 
summ(reg_female_fins, confint = T) 



#Estimating the mean and central tendency of BMD by screen time quartiles and EMM's
## Among Males
svyby(formula = ~DXDSTBMD, 
      by = ~factor(screen_quart), 
      design = subset(subpop_design, RIAGENDR =="Male" & fsdch_2cat == "High/Marginal Food Security"), # by sex male & FS
      FUN = svymean, 
      keep.var = TRUE)

svyby(formula = ~DXDSTBMD, 
      by = ~factor(screen_quart), 
      design = subset(subpop_design, RIAGENDR =="Male" & fsdch_2cat == "Low/Very Low Food Security"), # by sex male & FINS
      FUN = svymean, 
      keep.var = TRUE)


## Among Females
svyby(formula = ~DXDSTBMD, 
      by = ~factor(screen_quart), 
      design = subset(subpop_design, RIAGENDR =="Female" & fsdch_2cat == "High/Marginal Food Security"),# by sex female & FS
      FUN = svymean, 
      keep.var = TRUE)

svyby(formula = ~DXDSTBMD, 
      by = ~factor(screen_quart), 
      design = subset(subpop_design, RIAGENDR =="Female" & fsdch_2cat == "Low/Very Low Food Security"),# by sex female & FINS
      FUN = svymean, 
      keep.var = TRUE)


```

## Creation of Tables 4 and 5

### Among Males

```{r}

# 1. Distribution of Screen Quartile (Unweighted N, Weighted %)
fs_w_pct <- tbl_svysummary(
  data = subset(subpop_design, RIAGENDR == "Male" & fsdch_2cat=="High/Marginal Food Security"), 
  include = screen_quart, 
  missing = "no", 
  statistic = list(all_categorical() ~ "{n_unweighted} ({p}%)"), 
  label = list(screen_quart = "Screen Quartile")
) |> modify_header(stat_0 ~ "")


# 2. Mean BMD Table (Vertical Layout)
# -------------------------------------------------------------
# Calculate weighted means and SEs by screen_quart
bmd_by_q <- svyby(~DXDSTBMD, ~screen_quart, subset(subpop_design, RIAGENDR == "Male" & fsdch_2cat=="High/Marginal Food Security"), svymean, na.rm = TRUE)

# Create a named lookup vector formatted as "Mean (SE)"
bmd_lookup <- setNames(
  sprintf("%.3f (%.3f)", bmd_by_q$DXDSTBMD, bmd_by_q$se),
  as.character(bmd_by_q$screen_quart)
)

# Copy w_pct structure and map the Mean (SE) values onto the quartile levels
fs_mean <- fs_w_pct
fs_mean$table_body$stat_0 <- ifelse(
  fs_mean$table_body$row_type == "level",
  bmd_lookup[as.character(fs_mean$table_body$label)],
  NA_character_
)

fs_mean <- fs_mean |> 
  modify_header(stat_0 ~ "")

fs_mean

# 4. Adjusted Regression Table
t4 <- tbl_regression(
  reg_male_fs, 
  include = "factor(screen_quart)", 
  label = list("factor(screen_quart)" = "Screen Quartile"), 
  conf.level = 0.95, 
  conf.int = TRUE
)

# Harmonize internal variable names to match w_pct ("screen_quart")
t4$table_body$variable[t4$table_body$variable == "factor(screen_quart)"] <- "screen_quart"

strat_male_fs <- tbl_merge(
  tbls = list(fs_w_pct, fs_mean, t4),
  tab_spanner = c(
    "**Sample Size**",
    "**Mean BMD (SE)**",
    "**Stratified Model** <sup>3</sup>"
  )
) %>%
  modify_caption(
    caption = "Table 4. Associations beetween daily average screen time and TBLH BMD raw scores among male children aged 8- 17 y.o stratified by food security status (n = 2,044)"
  )

strat_male_fs


# 1. Distribution of Screen Quartile (Unweighted N, Weighted %)
fins_w_pct <- tbl_svysummary(
  data = subset(subpop_design, RIAGENDR == "Male" & fsdch_2cat=="Low/Very Low Food Security"), 
  include = screen_quart, 
  missing = "no", 
  statistic = list(all_categorical() ~ "{n_unweighted} ({p}%)"), 
  label = list(screen_quart = "Screen Quartile")
)

# 2. Mean BMD Table (Vertical Layout)
# -------------------------------------------------------------
# Calculate weighted means and SEs by screen_quart
bmd_by_q <- svyby(~DXDSTBMD, ~screen_quart, subset(subpop_design, RIAGENDR == "Male" & fsdch_2cat=="High/Marginal Food Security"), svymean, na.rm = TRUE)

# Create a named lookup vector formatted as "Mean (SE)"
bmd_lookup <- setNames(
  sprintf("%.3f (%.3f)", bmd_by_q$DXDSTBMD, bmd_by_q$se),
  as.character(bmd_by_q$screen_quart)
)

# Copy w_pct structure and map the Mean (SE) values onto the quartile levels
fins_mean <- fins_w_pct
fins_mean$table_body$stat_0 <- ifelse(
  fins_mean$table_body$row_type == "level",
  bmd_lookup[as.character(fins_mean$table_body$label)],
  NA_character_
)

fins_mean

# 4. Adjusted Regression Table
t5 <- tbl_regression(
  reg_male_fins, 
  exponentiate = FALSE, 
  include = "factor(screen_quart)", 
  label = list("factor(screen_quart)" = "Screen Quartile"), 
  conf.level = 0.95, 
  conf.int = TRUE
)

# Harmonize internal variable names to match w_pct ("screen_quart")
t5$table_body$variable[t5$table_body$variable == "factor(screen_quart)"] <- "screen_quart"

strat_male_fins <- tbl_merge(
  tbls = list(fins_w_pct, fins_mean, t5),
  tab_spanner = c(
    "**Sample Size**",
    "**Mean BMD (SE)**",
    "**Stratified Model** <sup>3</sup>"
  )
) 


strat_male_fins


# Stack them vertically
stacked_male_table <- tbl_stack(
  tbls = list(strat_male_fs, strat_male_fins),
  group_header = c("Food Secure Males", "Food Insecure Males") # Creates section headers
) %>%
  modify_caption(
    caption = "Table 4. Associations beetween daily average screen time and TBLH BMD raw scores among male children aged 8- 17 y.o stratified by food security status² (n = 2,044)"
  ) %>% 
  as_gt() %>%
  gt::tab_source_note(
    source_note = "² P-value for omnibus nested F-test (0.42)"
  ) |> 
  gt::tab_source_note(
    source_note = "³ Fully adjusted models were adjusted for age, race/ethnicity, height (in cm), household reference person's education level, and household income evaluated at the federal poverty level."
  )

stacked_male_table


# attempting to save the table 
gt::gtsave(stacked_male_table, filename = "Table4_Males.png") #save male as png


```

### Among Females

```{r}
# 1. Distribution of Screen Quartile (Unweighted N, Weighted %)
fs_w_pct <- tbl_svysummary(
  data = subset(subpop_design, RIAGENDR == "Female" & fsdch_2cat=="High/Marginal Food Security"), 
  include = screen_quart, 
  missing = "no", 
  statistic = list(all_categorical() ~ "{n_unweighted} ({p}%)"), 
  label = list(screen_quart = "Screen Quartile")
) |> modify_header(stat_0 ~ "")


# 2. Mean BMD Table (Vertical Layout)
# -------------------------------------------------------------
# Calculate weighted means and SEs by screen_quart
bmd_by_q <- svyby(~DXDSTBMD, ~screen_quart, subset(subpop_design, RIAGENDR == "Female" & fsdch_2cat=="High/Marginal Food Security"), svymean, na.rm = TRUE)

# Create a named lookup vector formatted as "Mean (SE)"
bmd_lookup <- setNames(
  sprintf("%.3f (%.3f)", bmd_by_q$DXDSTBMD, bmd_by_q$se),
  as.character(bmd_by_q$screen_quart)
)

# Copy w_pct structure and map the Mean (SE) values onto the quartile levels
fs_mean <- fs_w_pct
fs_mean$table_body$stat_0 <- ifelse(
  fs_mean$table_body$row_type == "level",
  bmd_lookup[as.character(fs_mean$table_body$label)],
  NA_character_
)

fs_mean <- fs_mean |> 
  modify_header(stat_0 ~ "")

fs_mean

# 4. Adjusted Regression Table
t4 <- tbl_regression(
  reg_female_fs, 
  include = "factor(screen_quart)", 
  label = list("factor(screen_quart)" = "Screen Quartile"), 
  conf.level = 0.95, 
  conf.int = TRUE
)

# Harmonize internal variable names to match w_pct ("screen_quart")
t4$table_body$variable[t4$table_body$variable == "factor(screen_quart)"] <- "screen_quart"

strat_female_fs <- tbl_merge(
  tbls = list(fs_w_pct, fs_mean, t4),
  tab_spanner = c(
    "**Sample Size**",
    "**Mean BMD (SE)**",
    "**Stratified Model** <sup>3</sup>"
  )
) %>%
  modify_caption(
    caption = "Table 5. Associations beetween daily average screen time and TBLH BMD raw scores among female children aged 8- 17 y.o stratified by food security status (n = 1,953)"
  )

strat_female_fs


# 1. Distribution of Screen Quartile (Unweighted N, Weighted %)
fins_w_pct <- tbl_svysummary(
  data = subset(subpop_design, RIAGENDR == "Female" & fsdch_2cat=="Low/Very Low Food Security"), 
  include = screen_quart, 
  missing = "no", 
  statistic = list(all_categorical() ~ "{n_unweighted} ({p}%)"), 
  label = list(screen_quart = "Screen Quartile")
)

# 2. Mean BMD Table (Vertical Layout)
# -------------------------------------------------------------
# Calculate weighted means and SEs by screen_quart
bmd_by_q <- svyby(~DXDSTBMD, ~screen_quart, subset(subpop_design, RIAGENDR == "Female" & fsdch_2cat=="High/Marginal Food Security"), svymean, na.rm = TRUE)

# Create a named lookup vector formatted as "Mean (SE)"
bmd_lookup <- setNames(
  sprintf("%.3f (%.3f)", bmd_by_q$DXDSTBMD, bmd_by_q$se),
  as.character(bmd_by_q$screen_quart)
)

# Copy w_pct structure and map the Mean (SE) values onto the quartile levels
fins_mean <- fins_w_pct
fins_mean$table_body$stat_0 <- ifelse(
  fins_mean$table_body$row_type == "level",
  bmd_lookup[as.character(fins_mean$table_body$label)],
  NA_character_
)

fins_mean

# 4. Adjusted Regression Table
t5 <- tbl_regression(
  reg_female_fins, 
  exponentiate = FALSE, 
  include = "factor(screen_quart)", 
  label = list("factor(screen_quart)" = "Screen Quartile"), 
  conf.level = 0.95, 
  conf.int = TRUE
)

# Harmonize internal variable names to match w_pct ("screen_quart")
t5$table_body$variable[t5$table_body$variable == "factor(screen_quart)"] <- "screen_quart"

strat_female_fins <- tbl_merge(
  tbls = list(fins_w_pct, fins_mean, t5),
  tab_spanner = c(
    "**Sample Size**",
    "**Mean BMD (SE)**",
    "**Stratified Model** <sup>3</sup>"
  )
) 


strat_female_fins


# Stack them vertically
stacked_female_table <- tbl_stack(
  tbls = list(strat_female_fs, strat_female_fins),
  group_header = c("Food Secure Females", "Food Insecure Females") # Creates section headers
) %>%
  modify_caption(
    caption = "Table 5. Associations beetween daily average screen time and TBLH BMD raw scores among female children aged 8- 17 y.o stratified by food security status² (n = 1,953)"
  ) %>% 
  as_gt() %>%
  gt::tab_source_note(
    source_note = "² P-value for omnibus nested F-test (0.97)"
  ) |> 
  gt::tab_source_note(
    source_note = "³ Fully adjusted models were adjusted for age, race/ethnicity, height (in cm), household reference person's education level, and household income evaluated at the federal poverty level."
  )

stacked_female_table


# attempting to save the table 
gt::gtsave(stacked_female_table, filename = "Table5_Females.png") #save male as png

```
