---
title: "BMD z-score Anabel"
subtitle: "EPI DATA 536"
editor: visual
date-modified: "today"
format: 
  html: 
    link-external-newwindow: true
    toc: true
    embed-resources: true
    self-contained: true
    page-layout: full
  pdf: default 
editor_options: 
  chunk_output_type: console
---

# Libraries & Dataset
```{r}
library(dplyr)
library(readr)
library(tidyr)
# install.packages("cdcanthro")
library(here)
library(janitor)

# I will be using the dataset from P3- "Analytic Sample_PartA"
# This dataset includes everybody from the 3 waves, not just our study population
nhanes_clean <- readRDS(here("data", "processed", "nhanes_bmd_pediatric_master.rds"))

# This dataset does have some variables created from prior z-score attempts...
# previously created variables will be checked.
```


# Step 1: create height-adjusted z-scores with package cdcanthro
```{r}
# creating an age indicator






nhanes_clean <- cdcanthro::cdcanthro(nhanes_clean,
age = RIDEXAGM,
ht = BMXHT)

nhanes_clean |> tabyl(RIDAGEYR, age_band)

nhanes_band <- nhanes_clean |>
filter(age_band >= 8, age_band <= 17) |>
group_by(age_band, sex) |>
summarize(
mean_group_bmd = mean(DXDSTBMD, na.rm = TRUE),
sd_group_bmd   = sd(DXDSTBMD, na.rm = TRUE),
.groups = "drop"
)

```










# Step 3: Calculate the unadjusted (for HAZ 😊) BMD-Z
nhanes_clean <- nhanes_clean |>
left_join(nhanes_band, by = c("age_band", "sex")) |>
mutate(
bmd_z_unadjusted = (DXDSTBMD - mean_group_bmd) / sd_group_bmd
)
# Step 4: create a linear model where the dependent variable is unadjusted BMD,
# independent variable is HAZ
bmd_model <- lm(bmd_z_unadjusted ~ haz0,
data = nhanes_clean)
# Step 5: use the model to predict the residual for each observation
#(whatever part of the child's BMD Z-score isn't explained by their height)
# and the SD for the residuals.
library(broom) # I think I can just use the augment function for this...
aug_model <- augment(bmd_model, data = nhanes_clean)
aug_model <- augment(bmd_model)
glimpse(aug_model)
?augment
resid_sd <- sd(aug_model$.resid, na.rm = TRUE)
# The final HAZ-adjusted BMD Z-score will then essentially be:
#individuals predicted residuals/the SD of residuals
nhanes_clean <- aug_model %>%
mutate(
tblh_bmd_z_adj = .resid / resid_sd
)
sum(!is.na(nhanes_clean$tblh_bmd_z_adj))
nhanes_clean |>
tabyl(age_band, tblh_bmd_z_adj)
remove(resid_sd)
remove(nhanes_clean)
# Step 1: create height-adjusted z-scores with package cdcanthro
nhanes_clean <- cdcanthro::cdcanthro(nhanes_clean,
age = RIDEXAGM,
ht = BMXHT)
# install.packages("cdcanthro")
#=========================================================================
# I will be using the dataset from P3- "Analytic Sample_PartA"
# Dataset includes everybody from the 3 waves, not just our study population
nhanes_clean <- readRDS(here("data",
"processed", "nhanes_bmd_pediatric_master.rds"))
# Step 1: create height-adjusted z-scores with package cdcanthro
nhanes_clean <- cdcanthro::cdcanthro(nhanes_clean,
age = RIDEXAGM,
ht = BMXHT)
# dataframe calculating the mean and SD for bands
nhanes_band <- nhanes_clean |>
filter(age_band >= 8, age_band <= 17) |>
group_by(age_band, sex) |>
summarize(
mean_group_bmd = mean(DXDSTBMD, na.rm = TRUE),
sd_group_bmd   = sd(DXDSTBMD, na.rm = TRUE),
.groups = "drop"
)
# Step 3: Calculate the unadjusted (for HAZ 😊) BMD-Z
nhanes_clean <- nhanes_clean |>
left_join(nhanes_band, by = c("age_band", "sex")) |>
mutate(
bmd_z_unadjusted = (DXDSTBMD - mean_group_bmd) / sd_group_bmd
)
# Step 4: create a linear model where the dependent variable is unadjusted BMD,
# independent variable is HAZ
bmd_model <- lm(bmd_z_unadjusted ~ haz0,
data = nhanes_clean)
aug_model <- augment(bmd_model)
glimpse(aug_model) # residuals but no SD of residuals...
View(aug_model)
aug_model <-aug_model |>
mutate(resid_sd = sd(.resid, na.rm = TRUE))
# The final HAZ-adjusted BMD Z-score will then essentially be:
# individuals predicted residuals/the SD of residuals
aug_model <- aug_model %>%
mutate(
tblh_bmd_z_adj = .resid / resid_sd
)
remove(aug_model)
nhanes_clean <- nhanes_clean |>
mutate(row_id = row_number())
remove(nhanes_clean)
# install.packages("cdcanthro")
#=========================================================================
# I will be using the dataset from P3- "Analytic Sample_PartA"
# Dataset includes everybody from the 3 waves, not just our study population
nhanes_clean <- readRDS(here("data",
"processed", "nhanes_bmd_pediatric_master.rds"))
# Step 1: create height-adjusted z-scores with package cdcanthro
nhanes_clean <- cdcanthro::cdcanthro(nhanes_clean,
age = RIDEXAGM,
ht = BMXHT)
# dataframe calculating the mean and SD for bands
nhanes_band <- nhanes_clean |>
filter(age_band >= 8, age_band <= 17) |>
group_by(age_band, sex) |>
summarize(
mean_group_bmd = mean(DXDSTBMD, na.rm = TRUE),
sd_group_bmd   = sd(DXDSTBMD, na.rm = TRUE),
.groups = "drop"
)
# Step 3: Calculate the unadjusted (for HAZ 😊) BMD-Z
nhanes_clean <- nhanes_clean |>
left_join(nhanes_band, by = c("age_band", "sex")) |>
mutate(
bmd_z_unadjusted = (DXDSTBMD - mean_group_bmd) / sd_group_bmd
)
# Step 4: create a linear model where the dependent variable is unadjusted BMD,
# independent variable is HAZ
bmd_model <- lm(bmd_z_unadjusted ~ haz0,
data = nhanes_clean)
nhanes_clean <- nhanes_clean |>
mutate(.rownames = row_number()) # creating a row number to left join later.
aug_model <- augment(bmd_model)
glimpse(aug_model) # residuals but no SD of residuals...
aug_model <-aug_model |>
mutate(resid_sd = sd(.resid, na.rm = TRUE)) # calc for SD of residuals
# The final HAZ-adjusted BMD Z-score will then essentially be:
# individuals predicted residuals/the SD of residuals
aug_model <- aug_model %>%
mutate(
tblh_bmd_z_adj = .resid / resid_sd
)
?left_join
# left join thse new values
aug_model_select <- aug_model |>
select(.rownames, .fitted, .resid, resid_sd, tblh_bmd_z_adj)
nhanes_bmd <- nhanes_clean |>
left_join(aug_model_select, by = ".rownames")
nhanes_clean <- nhanes_clean |>
mutate(.rownames = row_number()) |>
as.character() # creating a row number to left join later.
nhanes_bmd <- nhanes_clean |>
left_join(aug_model_select, by = ".rownames")
remove(nhanes_clean)
# install.packages("cdcanthro")
#=========================================================================
# I will be using the dataset from P3- "Analytic Sample_PartA"
# Dataset includes everybody from the 3 waves, not just our study population
nhanes_clean <- readRDS(here("data",
"processed", "nhanes_bmd_pediatric_master.rds"))
# Step 1: create height-adjusted z-scores with package cdcanthro
nhanes_clean <- cdcanthro::cdcanthro(nhanes_clean,
age = RIDEXAGM,
ht = BMXHT)
# dataframe calculating the mean and SD for bands
nhanes_band <- nhanes_clean |>
filter(age_band >= 8, age_band <= 17) |>
group_by(age_band, sex) |>
summarize(
mean_group_bmd = mean(DXDSTBMD, na.rm = TRUE),
sd_group_bmd   = sd(DXDSTBMD, na.rm = TRUE),
.groups = "drop"
)
# Step 3: Calculate the unadjusted (for HAZ 😊) BMD-Z
nhanes_clean <- nhanes_clean |>
left_join(nhanes_band, by = c("age_band", "sex")) |>
mutate(
bmd_z_unadjusted = (DXDSTBMD - mean_group_bmd) / sd_group_bmd
)
# Step 4: create a linear model where the dependent variable is unadjusted BMD,
# independent variable is HAZ
bmd_model <- lm(bmd_z_unadjusted ~ haz0,
data = nhanes_clean)
nhanes_clean <- nhanes_clean |>
mutate(.rownames = as.character(row_number())) # creating a row number to left join later.
nhanes_bmd <- nhanes_clean |>
left_join(aug_model_select, by = ".rownames")
nhanes_bmd |>
tabyl(age_band, tblh_bmd_z_adj)
View(nhanes_bmd)
# AI generated check
library(dplyr)
library(ggplot2)
# 1) Overall summary
nhanes %>%
summarize(
n = sum(!is.na(tblh_bmd_z_adj)),
mean_z = mean(tblh_bmd_z_adj, na.rm = TRUE),
sd_z   = sd(tblh_bmd_z_adj, na.rm = TRUE),
min_z  = min(tblh_bmd_z_adj, na.rm = TRUE),
max_z  = max(tblh_bmd_z_adj, na.rm = TRUE)
)
# 1) Overall summary
nhanes_bmd %>%
summarize(
n = sum(!is.na(tblh_bmd_z_adj)),
mean_z = mean(tblh_bmd_z_adj, na.rm = TRUE),
sd_z   = sd(tblh_bmd_z_adj, na.rm = TRUE),
min_z  = min(tblh_bmd_z_adj, na.rm = TRUE),
max_z  = max(tblh_bmd_z_adj, na.rm = TRUE)
)
# 1) Overall summary
nhanes_bmd %>%
summarize(
n = sum(!is.na(tblh_bmd_z_adj)),
mean_z = mean(tblh_bmd_z_adj, na.rm = TRUE),
sd_z   = sd(tblh_bmd_z_adj, na.rm = TRUE),
min_z  = min(tblh_bmd_z_adj, na.rm = TRUE),
max_z  = max(tblh_bmd_z_adj, na.rm = TRUE)
)
# 2) By age band and sex
nhanes_bmd %>%
group_by(age_band, sex) %>%
summarize(
n = sum(!is.na(tblh_bmd_z_adj)),
mean_z = mean(tblh_bmd_z_adj, na.rm = TRUE),
sd_z   = sd(tblh_bmd_z_adj, na.rm = TRUE),
.groups = "drop"
)
# 3) Z-score vs age
ggplot(nhanes, aes(x = age_years, y = tblh_bmd_z_adj, color = factor(sex))) +
geom_point(alpha = 0.3) +
geom_smooth(se = FALSE, method = "loess") +
geom_hline(yintercept = 0, linetype = "dashed") +
labs(x = "Age (years)", y = "Adjusted TBLH BMD z-score", color = "Sex")
# 3) Z-score vs age
ggplot(nhanes_bmd, aes(x = age_years, y = tblh_bmd_z_adj, color = factor(sex))) +
geom_point(alpha = 0.3) +
geom_smooth(se = FALSE, method = "loess") +
geom_hline(yintercept = 0, linetype = "dashed") +
labs(x = "Age (years)", y = "Adjusted TBLH BMD z-score", color = "Sex")
# 3) Z-score vs age
ggplot(nhanes_bmd, aes(x = RIDAGEYR, y = tblh_bmd_z_adj, color = factor(sex))) +
geom_point(alpha = 0.3) +
geom_smooth(se = FALSE, method = "loess") +
geom_hline(yintercept = 0, linetype = "dashed") +
labs(x = "Age (years)", y = "Adjusted TBLH BMD z-score", color = "Sex")
# 4) Z-score vs height
ggplot(nhanes_bmd, aes(x = BMXHT, y = tblh_bmd_z_adj, color = factor(sex))) +
geom_point(alpha = 0.3) +
geom_smooth(se = FALSE, method = "loess") +
geom_hline(yintercept = 0, linetype = "dashed") +
labs(x = "Height (cm)", y = "Adjusted TBLH BMD z-score", color = "Sex")
# 5) Histogram
ggplot(nhanes_bmd, aes(x = tblh_bmd_z_adj)) +
geom_histogram(bins = 30, color = "white") +
geom_vline(xintercept = 0, linetype = "dashed") +
labs(x = "Adjusted TBLH BMD z-score", y = "Count")
# 6) Q-Q plot
qqnorm(nhanes_bmd$tblh_bmd_z_adj)
qqline(nhanes_bmd$tblh_bmd_z_adj, col = "red")
# 7) If you still have the model object, residuals vs fitted
plot(aug_model$.fitted, aug_model$.resid,
xlab = "Fitted values", ylab = "Residuals")
# 7) If you still have the model object, residuals vs fitted
plot(aug_model$.fitted, aug_model$.resid,
xlab = "Fitted values", ylab = "Residuals") |>
abline(h = 0, col = "red", lty = 2)
# 1) Overall summary
nhanes_bmd %>%
summarize(
n = sum(!is.na(tblh_bmd_z_adj)),
mean_z = mean(tblh_bmd_z_adj, na.rm = TRUE),
sd_z   = sd(tblh_bmd_z_adj, na.rm = TRUE),
min_z  = min(tblh_bmd_z_adj, na.rm = TRUE),
max_z  = max(tblh_bmd_z_adj, na.rm = TRUE)
)
# 2) By age band and sex
nhanes_bmd %>%
group_by(age_band, sex) %>%
summarize(
n = sum(!is.na(tblh_bmd_z_adj)),
mean_z = mean(tblh_bmd_z_adj, na.rm = TRUE),
sd_z   = sd(tblh_bmd_z_adj, na.rm = TRUE),
.groups = "drop"
)
# 2) By age band and sex
print(nhanes_bmd %>%
group_by(age_band, sex) %>%
summarize(
n = sum(!is.na(tblh_bmd_z_adj)),
mean_z = mean(tblh_bmd_z_adj, na.rm = TRUE),
sd_z   = sd(tblh_bmd_z_adj, na.rm = TRUE),
.groups = "drop"
))
nhanes_bmd %>%
group_by(age_band, sex) %>%
summarize(
n = sum(!is.na(tblh_bmd_z_adj)),
mean_z = ifelse(n > 0, mean(tblh_bmd_z_adj, na.rm = TRUE), NA_real_),
sd_z   = ifelse(n > 1, sd(tblh_bmd_z_adj, na.rm = TRUE), NA_real_),
.groups = "drop"
)
print(nhanes_bmd %>%
group_by(age_band, sex) %>%
summarize(
n = sum(!is.na(tblh_bmd_z_adj)),
mean_z = ifelse(n > 0, mean(tblh_bmd_z_adj, na.rm = TRUE), NA_real_),
sd_z   = ifelse(n > 1, sd(tblh_bmd_z_adj, na.rm = TRUE), NA_real_),
.groups = "drop"
))
# Create a stable join key
nhanes_clean <- nhanes_clean |>
mutate(.rownames = as.character(row_number()))
# Fit/augment with the original data
aug_model <- augment(bmd_model, data = nhanes_clean)
# install.packages("cdcanthro")
#=========================================================================
# I will be using the dataset from P3- "Analytic Sample_PartA"
# Dataset includes everybody from the 3 waves, not just our study population
nhanes_clean <- readRDS(here("data",
"processed", "nhanes_bmd_pediatric_master.rds"))
# Step 1: create height-adjusted z-scores with package cdcanthro
nhanes_clean <- cdcanthro::cdcanthro(nhanes_clean,
age = RIDEXAGM,
ht = BMXHT)
# creating a dataframe calculating the mean and SD for bands
nhanes_band <- nhanes_clean |>
filter(age_band >= 8, age_band <= 17) |>
group_by(age_band, sex) |>
summarize(
mean_group_bmd = mean(DXDSTBMD, na.rm = TRUE),
sd_group_bmd   = sd(DXDSTBMD, na.rm = TRUE),
.groups = "drop"
)
# Step 3: Calculate the unadjusted (for HAZ 😊) BMD-Z
nhanes_clean <- nhanes_clean |>
left_join(nhanes_band, by = c("age_band", "sex")) |>
mutate(
bmd_z_unadjusted = (DXDSTBMD - mean_group_bmd) / sd_group_bmd
)
# Step 4: create a linear model where the dependent variable is unadjusted BMD,
# independent variable is HAZ
bmd_model <- lm(bmd_z_unadjusted ~ haz0,
data = nhanes_clean)
# Create a stable join key
nhanes_clean <- nhanes_clean |>
mutate(.rownames = as.character(row_number()))
# Fit/augment with the original data
aug_model <- augment(bmd_model, data = nhanes_clean)
rlang::last_trace()
remove(nhanes_clean)
# install.packages("cdcanthro")
#=========================================================================
# I will be using the dataset from P3- "Analytic Sample_PartA"
# Dataset includes everybody from the 3 waves, not just our study population
nhanes_clean <- readRDS(here("data",
"processed", "nhanes_bmd_pediatric_master.rds"))
# Step 1: create height-adjusted z-scores with package cdcanthro
nhanes_clean <- cdcanthro::cdcanthro(nhanes_clean,
age = RIDEXAGM,
ht = BMXHT)
# creating a dataframe calculating the mean and SD for bands
nhanes_band <- nhanes_clean |>
filter(age_band >= 8, age_band <= 17) |>
group_by(age_band, sex) |>
summarize(
mean_group_bmd = mean(DXDSTBMD, na.rm = TRUE),
sd_group_bmd   = sd(DXDSTBMD, na.rm = TRUE),
.groups = "drop"
)
# Step 3: Calculate the unadjusted (for HAZ 😊) BMD-Z
nhanes_clean <- nhanes_clean |>
left_join(nhanes_band, by = c("age_band", "sex")) |>
mutate(
bmd_z_unadjusted = (DXDSTBMD - mean_group_bmd) / sd_group_bmd
)
# Step 4: create a linear model where the dependent variable is unadjusted BMD,
# independent variable is HAZ
bmd_model <- lm(bmd_z_unadjusted ~ haz0,
data = nhanes_clean)
# I am going to try step 5 a different way
nhanes_clean <- nhanes_clean %>%
mutate(
bmd_predict   = predict(bmd_model, newdata = nhanes_clean),
bmd_resid  = DXDSTBMD - bmd_predict,
bmd_resid_sd = sd(bmd_resid, na.rm = TRUE),
tblh_bmd_z_adj = bmd_resid / bmd_resid_sd
)
View(nhanes_clean)
# 2) By age band and sex----> I am having issues with this, which tells me
# I may have done something wrong above.
print(nhanes_clean %>%
group_by(age_band, sex) %>%
summarize(
n = sum(!is.na(tblh_bmd_z_adj)),
mean_z = mean(tblh_bmd_z_adj, na.rm = TRUE),
sd_z   = sd(tblh_bmd_z_adj, na.rm = TRUE),
.groups = "drop"
))
install.packages("remotes")
remotes::install_github("jmcphers/rsrecovr")
rsrecovr::recovr()
savehistory("~/New_BMD-Project/recovered_202608042145/bmd_zscore_Anabel.Rhistory")

