---
title: "BMD z-score Anabel"
subtitle: "EPI DATA 536"
editor: visual
date-modified: "today"
format: 
  html: 
    link-external-newwindow: true
    toc: true
    embed-resources: true
    self-contained: true
    page-layout: full
  pdf: default 
editor_options: 
  chunk_output_type: console
---

# Libraries & Dataset

```{r}
library(dplyr)
library(readr)
library(tidyr)
# install.packages("cdcanthro")
library(here)
library(janitor)
library(cdcanthro)

# I will be using the dataset from P3- "Analytic Sample_PartA"
# This dataset includes everybody from the 3 waves, not just our study population
nhanes_clean <- readRDS(here("data", "processed", "nhanes_bmd_pediatric_master.rds"))

# This dataset does have some variables created from prior z-score attempts...
# previously created variables will be checked, if used.
```

# Step 1: create height-adjusted z-scores with package cdcanthro

```{r}
# creating an age indicator
nhanes_clean <- nhanes_clean |> 
  mutate(
    age_instudy = case_when(
      RIDAGEYR >= 0 & RIDAGEYR <= 7 ~ "0", 
      RIDAGEYR >= 8 & RIDAGEYR <=17 ~ "1", 
      RIDAGEYR > 17 ~ "0", 
      TRUE ~ NA_character_) |> 
  factor(levels = c("0", "1")))

nhanes_clean |> 
tabyl(RIDAGEYR, age_instudy) # checking to make sure indicator is coded correct

# convert age in years to age in months.
# there is already a variable in NHANES that has age in months up to 19 y.o
# it is called RIDEXAGM, however some people in our study pop are missing this var,

sum(is.na(nhanes_clean$RIDEXAGM))
sum(is.na(nhanes_clean$BMXHT))
sum(is.na(nhanes_clean$RIDAGEYR))

# so ill just create a var with RIDAGEYR since missing = 0
nhanes_clean <- nhanes_clean |> 
  mutate(age_months = RIDAGEYR*12 + 6) 
sum(is.na(nhanes_clean$age_months)) # check, no missing values. 

# use cdcanthro() to calculate haz scores. Because cdcanthro is limited for a certain age range, and we’re not dropping anyone from the original NHANES in order to preserve the weighting, try:

haz_results <- cdcanthro(data=subset(nhanes_clean, age_instudy==1), 
                         age=age_months, ht = BMXHT)

# because we have a prior variable called "haz" (from prior attempts), the newly created 
# variable is called "haz0". 

class(haz_results) # check to make sure it is d.f

# left join back to the large data frame
nhanes_full <- nhanes_clean %>% left_join(haz_results, by= "SEQN")
# there are duplicate columns in our df now, demarcated by ".x" or ".y"
# (no prob...)

# checking to see how many HAZ we are missing in our study pop
nhanes_full %>%
  group_by(age_instudy.x) %>%
  summarise(na_count = sum(is.na(haz0)),
            non_na_count = sum(!is.na(haz0))) 
# we have 272 people in our study population who are missing a HAZ score. 

nhanes_full %>%
  group_by(age_instudy.x) %>%
  summarise(na_count = sum(is.na(BMXHT.x)),
            non_na_count = sum(!is.na(BMXHT.x))) 

# because we have 272 people who are missing a height... just checking

```

# Step 2: Run Survey design using svydesign() on the updated NHANES data with the haz scores

```{r}
# Temporary survey design with the new 'haz0' score
library(survey)
design <- svydesign(
  id = ~SDMVPSU.x,
  strata = ~SDMVSTRA.x,
  weights = ~WTMEC6YR.x,
  nest = TRUE,
  data = nhanes_full) 

```

# Step 3: Calculate weighted group-specific mean and variance (to get SD) for BMD (age x sex only)

```{r}
# calculating weighted mean 
group_mean_stats <- svyby(
  formula = ~DXDSTBMD.x,
  by = ~RIDAGEYR.x + RIAGENDR.x,
  design = design,
  FUN = svymean,
  na.rm = TRUE,
  keep.var = TRUE
) |>
  as.data.frame()

names(group_mean_stats)[3] <- "group_mean" # changing name for [3] of the list

group_mean_stats <- group_mean_stats |>
  select(RIDAGEYR.x, RIAGENDR.x, group_mean) # removes (se) column

group_mean_stats # columns exist, although unclear why some ages have means of 0.
# however, our study pop of interest has means. 

# calculating weighted SD
group_var_stats <- svyby(
  formula = ~DXDSTBMD.x,
  by = ~RIDAGEYR.x + RIAGENDR.x,
  design = design,
  FUN = svyvar,
  na.rm = TRUE,
  keep.var = TRUE
) |>
  as.data.frame()

names(group_var_stats)[3] <- "group_var" # changing name for [3] of the list

group_var_stats <- group_var_stats |>
  select(RIDAGEYR.x, RIAGENDR.x, group_var)

group_var_stats # columns exist, hard to tell if they look fully OK.


# Merge means and variance and calculate SDs by taking the square root of the variance:
  group_stats <- left_join(group_mean_stats, group_var_stats, by = c("RIDAGEYR.x", "RIAGENDR.x")) %>% mutate(group_sd = sqrt(group_var))

group_stats # I have all the columns (mean, var, and sd)
  
```

# Step 4: Merge back to the big NHANES dataset and compute unadjusted BMD z-scores

```{r}
# merging the df above to the complete NHANES dataset
nhanes_merge <- left_join(nhanes_full, group_stats, 
                          by = c("RIDAGEYR.x", 
                                "RIAGENDR.x")) |> 

  mutate(z_unadj = 
           (DXDSTBMD.x - group_mean) / group_sd) # computing unadjusted BMD 
#and creating new column var 


nhanes_merge %>%
  group_by(age_instudy.x) %>%
  summarise(na_count = sum(is.na(z_unadj)),
            non_na_count = sum(!is.na(z_unadj))) 

# missing n = 1062 in our age_instudy pop who are missing a z-unadj score. 
# I'd like to check this and see if the n missing should be this large.

```

# Step 5: Calculate HAZ-adjusted BMD z-scores

```{r}
# Create a filter indicator for complete unadjusted z-scores (z_unadj variable) (yes/no)

nhanes_merge <- nhanes_merge |> 
  mutate(
    z_unadj_cat = if_else(is.na(z_unadj), "no", "yes"))

# what is this indicator variable for? 

# check    
table(is.na(nhanes_merge$z_unadj), nhanes_merge$z_unadj_cat)
# unsure if this indicator variable worked properly


#Run your survey design again, applying the indicator for complete unadjusted z scores above.
# Let’s call this survey design “design” still.
design <- svydesign(
  id = ~SDMVPSU.x,
  strata = ~SDMVSTRA.x,
  weights = ~WTMEC6YR.x,
  nest = TRUE,
  data = nhanes_merge) 

# Run a linear model: 
haz_model <- svyglm(z_unadj ~ haz0, design = design) 
# I ran the linear model with the the continuous "z_unadj" rather than indicator.

library(jtools)
summ(haz_model)
# for some reason the haz model has 4504 observations which is causing issues with the code lines below. 


# Now, in your NHANES data, create new variables for individual predicted z scores and individual residual from the haz_model above. The residual is the unadjusted z scores minus the predicted

nhanes_merge <- nhanes_merge %>% mutate(
  z_pred = fitted(haz_model),
  z_resid = (z_unadj - z_pred))


# Run your survey design again since you just created new variables – for simplicity I still call it “design” for now.


# Get the variance of the residual z scores to calculate their SD
resid_var <- svyvar(~z_resid, design = design)

resid_sd <- sqrt(as.numeric(resid_var))

# Now, obtain the HAZ-adjusted BMD z scores:
nhanes <- nhanes %>% mutate(z_haz_adj = z_resid / resid_sd)

```
